Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline.

### 2017-04-22 draft/ms_v4.tex
Update until Section 6.1

    %

    These challenges are not merely technical details; their resolution is crucial
    to achieving key scientific goals.
    For instance, weak gravitational lensing (WL; \citealt{Kaiser1993, Bartelmann2001})
    is a powerful tool for measuring the large-scale distribution of dark matter.
    However, WL measurements critically depend on our ability to measure the shape and photometric redshifts of background galaxies with high precision. Photometric redshifts (e.g., \citealt{Benitez2000, Bolzonella2000,
    Ilbert2009}) are fundamentally important for studying the evolution of galaxies,
    and the ``drop-out'' method is critical for selecting high-redshift galaxies
    (e.g., \citealt{Steidel1996}).
    However, both of those methods reply on the availability of accurate multi-band photometric
    measurements. With increasingly large surveys,  and with more stringent requirements on data quality,  quality control also becomes a pressing issue.

@@

 \alexie{Edit here following this format}. In Section 1 we ... in section 2 we ....

    Below, we briefly introduce the HSC survey and the current status of data
    reduction (Section \ref{sec:ssp}), explain \synpipe{} design and implementation
    (Section \ref{sec:synpipe}), and demonstrate \synpipe{} usage via straightforward
    tests (Section \ref{sec:test}) that focus on the general photometric quality for
    both synthetic stars and galaxies (Section \ref{sec:result}).
    We then summarize the work and discuss future developments in
    Section \ref{sec:summary}.

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Flowchart
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_flowchart_new}
    \end{center}
    \caption{
        Illustration of the main steps involved in using \synpipe{}.
        Gray boxes indicate required inputs at different stages.
        The blue box identifies the \texttt{addFakes.py} step, when \synpipe{} injects synthetic objects into single-frame images. Below the blue box, we show an HSC image before and after the insertion of synthetic galaxies. The positions of synthetic  objects are highlighted with green circles. Red boxes depict the image coadding and multi-band measurement steps
        using \texttt{stack.py} and \texttt{multiBand.py}. At the bottom left, we show a \coadd{} image which contains synthetic galaxies. On the right hand side, we show the spatial relation
        between \texttt{Tracts}, \texttt{Patches}, and \texttt{Visits}. The xxxxxxx colored box corresponds to one \texttt{Tract} which has an area of about xx deg$^2$. Large colored circles are \texttt{Visits} (also commonly known as "pointings") with small rectangles representing CCDs. \texttt{Patches} are represented by black dashed lines. One \texttt{Tract} typically contains xx  \texttt{Patches}.}
    \label{fig:flowchart}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

        \item \textbf{Single-\texttt{Visit~}Processing}:
            Each individual exposure is called a \term{Visit} and is assigned
            an even integer.
            After initial data screening (considering the background level, seeing, and transparency
            Furusawa\etal in prep.), \hscpipe{} detrends these data (subtraction
            of overscan, bias, and dark; flatfielding), \alexie{xxxxx} generates variance and mask
            images, subtracts the background, and provides initial astrometric and
            photometric calibrations. \synpipe{} uses the spatially varying PSF models, photometric zeropoints, and the World Coordinate System (WCS; \citealt{WCS1, WCS2})
            corrected for optical distortion provided by \hscpipe{}.
        \item \textbf{Multi-\visit{} Processing}:
            After single-\visit{} processing, \hscpipe{} warps and mosaics the
            reduced CCD images into much deeper \coadd{} images while improving
            the astrometric and photometric calibrations via processes similar to
            the \"{u}ber-calibration in the Sloan Digital Sky Survey (SDSS)
            \citep{Padmanabhan2008}.
            \textbf{\synpipe{} will use these improved calibrations.}
            \hscpipe{} organizes these \coadd{} images into equiareal rectangular
            regions, or \tracts{}, which are predefined as iso-latitude tessellations.
            One \tract{} covers approximately $1.7\times 1.7$ degrees$^2$ and
            adjacent \tracts{} overlap each other by ${\sim}1$\amin{}.
            Each \tract{} is further divided into $9\times9$ \texttt{Patches}.
            A \texttt{Patch} is a $4200\times4200$ pixels rectangular region, and
            adjacent \texttt{Patches} overlap each other by 100 pixels.   
        \item \textbf{Multi-band Measurements}:
            To achieve consistent photometry across all filters, \hscpipe{} first
            detects and deblends objects on \coadd{} images in each band
            independently (the \unforced{} measurements).
            The collection of above-threshold pixels for each object is referred to as
            a \texttt{footprint}.
            \hscpipe{} merges the \texttt{footprints} and flux peaks in the different
            bands and then selects a reference band for each object based on its $S/N$ in each band (usually this corresponds to the $i$-band).  Centroids, shapes, and other non-amplitude parameters are then fixed to  the values from the reference band. \hscpipe{} then performs \forced{} photometry using these fixed quantities. The goal of the  \forced{} photometry step is to generate accurate colors.
        \item \textbf{HSC Photometry}:
            The multi-band catalogs generated by \hscpipe{} contain various types of photometry measurements (see \citealt{HSCDR1} for details). Here, we focus on characterizing the the \forced{} \texttt{psf} and \cmodel{}
            photometry. The
            \texttt{psf} photometry should provide the most appropriate magnitudes and
            colors for point sources. The  \cmodel{} should be used for galaxies. The HSC  \cmodel{} algorithm is described in Bosch\etal, in prep. and is an  improved version of the SDSS \cmodel{} magnitude (\citealt{Lupton2001, Abazajian2004}). The HSC  \cmodel{} fits objects using both a de Vaucouleurs and an exponential profiles taking into account the smearing by the PSF.  \cmodel{} then linearly combines the two results to find the best fit to the galaxy image. \cmodel{} magnitudes are designed to yield accurate fluxes and colors for galaxies. Star--galaxy separation is performed based on
            the difference between \texttt{psf} and \cmodel{}{} magnitudes.
    \end{enumerate}

    DR1 data quality has been vetted in \citet{HSCDR1} by cross-checking
    with the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 1
    imaging survey (\citealt{Schlafly2012, Tonry2012, Magnier2013}) to examine the
    behavior of the stellar sequence.
    Generally speaking, the PSF photometry is accurate at the 1--2\% level \alexie{only for bright stars}, and the
    astrometry is accurate to ${\sim}10$ and 40 max internally and externally.

@@

    \alexie{Can we use bullet points to highlight the different steps? This paragraph is really hard to read.}

    With the input catalog and test data information in hand, \synpipe{}injects synthetic
    objects into single-exposure images (step \texttt{addFakes.py}; see the middle panel
    of Fig \ref{fig:flowchart}).
    \synpipe{} goes through all the individual CCD images that belong to each \visit{}
    and decides which synthetic objects from the input catalog need to be injected.
    \synpipe{} uses the \textit{initial astrometric calibration} of each single
    \visit{} to convert the input coordinates into locations on the CCD image. For each synthetic object, \synpipe{} uses the  reconstructed PSF for each exposure.  The photometric zero point from the single-\visit{} calibration is used to convert input magnitudes of synthetic objects into fluxes. For point sources, \synpipe{} generates a rectangular cutout image of the PSF model
    with appropriate size and correct total flux. For galaxies described by single- or double-\ser{} models, \synpipe{} passes the
    input \ser{} index and effective radius to the \galsim{}\texttt{.Sersic} function to
    create a \ser{} component.
    After stretching and rotating the galaxy to the expected axis ratio and position
    angle via the \texttt{shear} and \texttt{rotate} methods, \synpipe{} assigns the
    correct flux to this component.
    For a double-\ser{} model, \synpipe{} uses the \galsim{}\texttt{.Add} method to
    combine two \ser{} components. For  parametric models from the GREAT3 catalog, \synpipe{} uses the
    \texttt{COSMOSCatalog.makeGalaxy} method to generate models.  To inject  real HST galaxy images,  \synpipe{} calls the \galsim{}\texttt{.RealGalaxy} method.
    If necessary, an additional shear ($g_1$ and $g_2$) can be applied to the model at
    this point.
    Then, \synpipe{} passes the reconstructed PSF image into a
    \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy model
    using \galsim{}\texttt{.Convolve}.
    After this, \synpipe{} uses the \galsim{}\texttt{.drawImage} method to create the
    image of the simulated galaxy.
    A component with a high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires a
    large image size to cover all of its flux, so \synpipe{} allows the user to
    truncate the model at a given radius (N$\times R_{\mathrm{e}}$).
    \synpipe{} will also give warning information when a component with $
    n_{\mathrm{Ser}} > 6$ is encountered because it takes much longer to achieve
    accurate PSF convolution in such a model; and the result is, therefore,
    not a realistic galaxy model.

    After \synpipe{} creates the image of the simulated galaxy, it then shifts the
    image according to is location on the CCD on a pixel-by-pixel basis;
    \synpipe{} also crops the images when necessary.
    When that task is complete, \synpipe{} adds appropriate noise to the image based
    on the pixels it covers and the calibration information of the detector, and creates
    a variance image that reflects the influence of the synthetic object.
    Since each HSC CCD consists of 4 amplifiers, each with different characteristics,
    \synpipe{} carefully creates a corresponding gain map to provide accurate noise
    level.
    In the final step, \synpipe{} adds the noise-added image and the variance map to the
    original CCD data while also adding a new \texttt{FAKE} mask bit to the mask plane.
    The middle panel of Fig \ref{fig:flowchart} depicts the result of the
    \texttt{addFakes.py} step by showing an example CCD image before and after the
    synthetic objects are injected.
%% ------------------------------------------------------------------------------------ %%

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic stars and the $log(S/N)$ measured by \hscpipe{} PSF
        photometry.
        Hexagonal binned density plots and contours show the distributions for
        stars in both the good- and bad-seeing \tracts{}. Highly blended stars are highlighted using scatter plots.
        The gray dashed line marks $S/N = 5$ which is the detection threshold used by \hscpipe{}.  Our 5$\sigma$  point source detection limit  is \alexie{xxxx} however we do not reach these magnitudes with our current synthetic stellar input catalog.
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    We randomly inject ${\sim}100,000$ stars into each \tract{} in all
    five bands. Approximately ${\sim}3$--$4$\% of the stars are located within 2 pixels of the centroids
    of real objects; we remove those stars from the sample to avoid confusion in the comparisons. \alexie{xxxxxx}
 %   After this step, ${\sim}1$--$3$\% of the remaining synthetic stars still have no match in the \synpipe{} outputs.
 %   They could either be below the detection threshold or ambiguously blended with real objects \alexie{said you identified those}.
 %   The input $i$-band magnitudes of these unmatched stars are on average ${\sim}0.9$
 %   mag fainter than the matched ones.
 %   The fraction of unmatched stars is slightly higher in the bad seeing \tract{}; this result is consistent with our expectation that magnitude depends on seeing conditions,
 %   although the difference is not very significant \alexie{cut?}.
    For matched stars, we select the primary detections
    (\texttt{detect.is-primary=True}) that have good photometric quality
    (see Appendix \ref{app:qc} for details).
    This gives us ${\sim}83,000$ stars in each \tract{}.

>>>>

    %
    These challenges are not merely technical details; their resolution is crucial
    to achieving key scientific goals.
    For instance, weak gravitational lensing (WL; \citealt{Kaiser1993, Bartelmann2001})
    is a powerful tool for measuring the large-scale distribution of dark matter.
    However, WL measurements critically depend on our ability to measure the shape and photometric redshifts of background galaxies with high precision. Photometric redshifts (e.g., \citealt{Benitez2000, Bolzonella2000,
    Ilbert2009}) are fundamentally important for studying the evolution of galaxies,
    and the ``drop-out'' method is critical for selecting high-redshift galaxies
    (e.g., \citealt{Steidel1996}).
    However, both of those methods reply on the availability of accurate multi-band photometric
    measurements. With increasingly large surveys,  and with more stringent requirements on data quality,  quality control also becomes a pressing issue.

@@

    In Section \ref{sec:ssp}, we briefly introduce the HSC survey and the current status
    of data reductionin. 
    In Section \ref{sec:synpipe}, we explain \synpipe{} design and implementation.
    In Section \ref{sec:test}, we demonstrate \synpipe{} usage via straightforward
    tests, we show the main results for the general photometric quality of synthetic 
    stars and galaxies in Section \ref{sec:result}.
    We then summarize the work and discuss future developments in Section 
    \ref{sec:summary}.

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Flowchart
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_flowchart_new}
    \end{center}
    \caption{
        Illustration of the main steps involved in using \synpipe{}.
        Gray boxes indicate required inputs at different stages.
        The blue box identifies the \texttt{addFakes.py} step, when \synpipe{} injects synthetic objects into single-frame images. Below the blue box, we show an HSC image before and after the insertion of synthetic galaxies. The positions of synthetic  objects are highlighted with green circles. Red boxes depict the image coadding and multi-band measurement steps
        using \texttt{stack.py} and \texttt{multiBand.py}. At the bottom left, we show a \coadd{} image which contains synthetic galaxies. On the right hand side, we show the spatial relation
        between \texttt{Tracts}, \texttt{Patches}, and \texttt{Visits}. 
        The red colored box corresponds to one \texttt{Tract} which has an area of 
        about 1.5 deg$^2$. 
        Large colored circles are \texttt{Visits} (also commonly known as "pointings") 
        with small rectangles representing CCDs. 
        \texttt{Patches} are represented by black dashed lines. 
        One \texttt{Tract} typically contains 81 \texttt{Patches}.}
    \label{fig:flowchart}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

        \item \textbf{Single-\texttt{Visit~}Processing}:
            Each individual exposure is called a \term{Visit} and is assigned
            an even integer.
            After initial data screening (considering the background level, seeing, 
            and transparency Furusawa\etal in prep.), \hscpipe{} subtracts overscan, 
            bias, and dark frames, and performs flatfielding to the single CCD images.
            During this step, \hscpipe{} also generates variance and mask images, 
            subtracts the background, and provides initial astrometric and photometric
            calibrations. 
            \synpipe{} uses the spatially varying PSF models, photometric zeropoints, 
            and the World Coordinate System (WCS; \citealt{WCS1, WCS2})
            corrected for optical distortion provided by \hscpipe{}.

        \item \textbf{Multi-\visit{} Processing}:
            After single-\visit{} processing, \hscpipe{} warps and mosaics the
            reduced CCD images into much deeper \coadd{} images while improving
            the astrometric and photometric calibrations via processes similar to
            the \"{u}ber-calibration in the Sloan Digital Sky Survey (SDSS)
            \citep{Padmanabhan2008}.
            \textbf{\synpipe{} will use these improved calibrations.}
            \hscpipe{} organizes these \coadd{} images into equiareal rectangular
            regions, or \tracts{}, which are predefined as iso-latitude tessellations.
            One \tract{} covers approximately $1.7\times 1.7$ degrees$^2$ and
            adjacent \tracts{} overlap each other by ${\sim}1$\amin{}.
            Each \tract{} is further divided into $9\times9$ \texttt{Patches}.
            A \texttt{Patch} is a $4200\times4200$ pixels rectangular region, and
            adjacent \texttt{Patches} overlap each other by 100 pixels.

        \item \textbf{Multi-band Measurements}:
            To achieve consistent photometry across all filters, \hscpipe{} first
            detects and deblends objects on \coadd{} images in each band
            independently (the \unforced{} measurements).
            The collection of above-threshold pixels for each object is referred to as
            a \texttt{footprint}.
            \hscpipe{} merges the \texttt{footprints} and flux peaks in the different
            bands and then selects a reference band for each object based on its $S/N$ in each band (usually this corresponds to the $i$-band).  Centroids, shapes, and other non-amplitude parameters are then fixed to  the values from the reference band. \hscpipe{} then performs \forced{} photometry using these fixed quantities. The goal of the  \forced{} photometry step is to generate accurate colors.

        \item \textbf{HSC Photometry}:
            The multi-band catalogs generated by \hscpipe{} contain various types of photometry measurements (see \citealt{HSCDR1} for details). Here, we focus on characterizing the the \forced{} \texttt{psf} and \cmodel{}
            photometry. The
            \texttt{psf} photometry should provide the most appropriate magnitudes and
            colors for point sources. The  \cmodel{} should be used for galaxies. The HSC  \cmodel{} algorithm is described in Bosch\etal, in prep. and is an  improved version of the SDSS \cmodel{} magnitude (\citealt{Lupton2001, Abazajian2004}). The HSC  \cmodel{} fits objects using both a de Vaucouleurs and an exponential profiles taking into account the smearing by the PSF.  \cmodel{} then linearly combines the two results to find the best fit to the galaxy image. \cmodel{} magnitudes are designed to yield accurate fluxes and colors for galaxies. Star--galaxy separation is performed based on
            the difference between \texttt{psf} and \cmodel{}{} magnitudes.

    \end{enumerate}
%% ------------------------------------------------------------------------------------ %%

    DR1 data quality has been vetted in \citet{HSCDR1} by cross-checking
    with the Panoramic Survey Telescope and Rapid Response System (Pan-STARRS) 1
    imaging survey (\citealt{Schlafly2012, Tonry2012, Magnier2013}) to examine the
    behavior of the stellar sequence.
    Generally speaking, the PSF photometry is accurate at the 1--2\% level for bright 
    stars, and the astrometry is accurate to ${\sim}10$ and 40 max internally and 
    externally.

@@

        \item With the input catalog and test data information in hand, \synpipe{} 
            injects synthetic objects into single-exposure images (step
            \texttt{addFakes.py}; see the middle panel of Fig \ref{fig:flowchart}).
            \synpipe{} goes through all the individual CCD images that belong to 
            each \visit{} and decides which synthetic objects from the input catalog 
            need to be injected. 
            \synpipe{} uses the \textit{initial astrometric calibration} of each 
            single \visit{} to convert the input coordinates into locations on the 
            CCD image. 
            For each synthetic object, \synpipe{} uses the reconstructed PSF for each 
            exposure.  
            The photometric zero point from the single-\visit{} calibration is used
            to convert input magnitudes of synthetic objects into fluxes.

        \item With the help of \galsim{} module, \synpipe{} simulate the images of
            synthetic objects.
            \begin{itemize}

            \item For point sources, \synpipe{} generates a rectangular cutout image 
                of the PSF model with appropriate size and correct total flux.

            \item For galaxies described by single- or double-\ser{} models, 
                \synpipe{} passes the input \ser{} index and effective radius to 
                the \galsim{}\texttt{.Sersic} function to create a \ser{} component.
                After stretching and rotating the galaxy to the expected axis ratio 
                and position angle via the \texttt{shear} and \texttt{rotate} methods,
                \synpipe{} assigns the correct flux to this component.

            \item For a double-\ser{} model, \synpipe{} uses the \galsim{}\texttt{.Add}
                method to combine two \ser{} components.

            \item For parametric models from the GREAT3 catalog, \synpipe{} uses the
                \texttt{COSMOSCatalog.makeGalaxy} method to generate models.

            \item To inject real HST galaxy images, \synpipe{} calls the \galsim{}    
                \texttt{.RealGalaxy} method.

        If necessary, an additional shear ($g_1$ and $g_2$) can be applied to the 
        model at this point.
        Then, \synpipe{} passes the reconstructed PSF image into a
        \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy 
        model using \galsim{}\texttt{.Convolve}.
        After this, \synpipe{} uses the \galsim{}\texttt{.drawImage} method to create 
        the image of the simulated galaxy.
        A component with a high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires 
        a large image size to cover all of its flux, so \synpipe{} allows the user 
        to truncate the model at a given radius (N$\times R_{\mathrm{e}}$).
        \synpipe{} will also give warning information when a component with $
        n_{\mathrm{Ser}} > 6$ is encountered because it takes much longer to achieve
        accurate PSF convolution in such a model; and the result is, therefore,
        not a realistic galaxy model.

        \item After \synpipe{} creates the image of the simulated galaxy, it then 
            shifts the image according to is location on the CCD on a pixel-by-pixel 
            basis; \synpipe{} also crops the images when necessary.
            When that task is complete, \synpipe{} adds appropriate noise to the 
            image based on the pixels it covers and the calibration information of 
            the detector, and creates a variance image that reflects the influence 
            of the synthetic object.
            Since each HSC CCD consists of 4 amplifiers, each with different 
            characteristics, \synpipe{} carefully creates a corresponding gain map 
            to provide accurate noise level.

        \item In the final step, \synpipe{} adds the noise-added image and the 
            variance map to the original CCD data while also adding a new 
            \texttt{FAKE} mask bit to the mask plane.
            The middle panel of Fig \ref{fig:flowchart} depicts the result of the
            \texttt{addFakes.py} step by showing an example CCD image before and 
            after the synthetic objects are injected.

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic stars and the $log(S/N)$ measured by \hscpipe{} PSF
        photometry.
        Hexagonal binned density plots and contours show the distributions for
        stars in both the good- and bad-seeing \tracts{}. 
        Highly blended stars are highlighted using scatter plots.
        The gray dashed line marks $S/N = 5$ which is the detection threshold used 
        by \hscpipe{}.  
        Our 5$\sigma$ point source detection limit is ${\sim}26.5$ mag in $i$-band,
        however we do not reach these magnitudes with our current synthetic stellar 
        input catalog.
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    We randomly inject ${\sim}100,000$ stars into each \tract{} in all
    five bands. Approximately ${\sim}3$--$4$\% of the stars are located within 2 pixels of the centroids
    of real objects; we remove those stars from the sample to avoid confusion in the comparisons.
     %   After this step, ${\sim}1$--$3$\% of the remaining synthetic stars still have no match in the \synpipe{} outputs.
 %   They could either be below the detection threshold or ambiguously blended with real objects \alexie{said you identified those}.
 %   The input $i$-band magnitudes of these unmatched stars are on average ${\sim}0.9$
 %   mag fainter than the matched ones.
 %   The fraction of unmatched stars is slightly higher in the bad seeing \tract{}; this result is consistent with our expectation that magnitude depends on seeing conditions,
 %   although the difference is not very significant \alexie{cut?}.
    For matched stars, we select the primary detections
    (\texttt{detect.is-primary=True}) that have good photometric quality
    (see Appendix \ref{app:qc} for details).
    This gives us ${\sim}83,000$ stars in each \tract{}.
